---
title: 'Design: Evaluation of Precision'
author: "Kai Gu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reference

> CLSI EP05-A3, "Evaluation of Precision of Quantitative Measurement Procedures"

### Design

This evaluation of precision is not common used in IVD trials, but it is necessary 
to include this process in QC procedure for end-users laboratories for verifying 
repeatability and within-laboratory precision.

So somewhere it's not very important but essential to operation, and communicating 
with laboratories users. For example, they will ask what's the precision (%CV and SD) 
of the within-run, between-run and between day.

### Implement

For now I trend to use `VCA` package instead of re-writing the nested ANOVA model to
evaluate the precision. For one side, it can save some time to develop and reduce the 
error rate although I have compared with `VCA` package to ensure my code is consistent 
with it. For another, this `VCA` package was developed by Roche Team, and is the same 
as the `mcr` package. And I have used these two R packages a lot, so I can guarantee
the accuracy of them.

```{r}
library(VCA)
data(dataEP05A2_2)
res <- anovaVCA(y~day/run, dataEP05A2_2)
res

VCAinference(res)
```

However the drawback to this is that I have to import the entire `VCA` package, 
like the same practice to `mcr` package. Actually I don't like to do that, because 
it will lead `mcradds` package appear too bloated since I only require a partial 
analysis of `VCA`.

So if needed, I will re-write it so that in can be more focused on out target in 
a later release update.
