---
title: 'Design: Class structure'
author: "Kai Gu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This vignette explores what could be a good class inheritance design in the `mcradds` package.

## Objectives

This package is a complement to `mcr`, so it means we can select the same class structure as `mcr`package. Or we just want to fast develop, so S3 class could be a choice.
- S4 class, similar to `mcr` package.
- S3 class, fast to develop.

We need to deal with two data categories, quantitative and qualitative data such as:
- `c(1, 2, 3)` as the quantitative value of one measurement method.
- `c("Positive", "Negative")` as the qualitative value of one measurement method.

Below methods will be implemented in above two data categories.
- quantitative data:
  - Descriptive statistics, frequencies.
  - Regression methods, rely on `mcr` package.
  - Bland-Altman for quantitative data, refer to CLSI EP09.
  - Outlier detection for quantitative data, refer to CLSI EP09-A2/A3.
  - Medical Decision Level, rely on `mcr` package, and summarize the results.
  - Reference Range/Interval, refer to CLSI EP28-A3, or NMPA guideline.
  - ROC/AUC, wraps the `pROC` package.
  - Correlation
- qualitative data:
  - Descriptive statistics, proportion/ratios.
  - Diagnostic Accuracy Criteria for quantitative data, refers to CLSI EP12-A2.
  - ROC/AUC, wraps the `pROC` package.

Figures can be displayed by some simple functions, like `boxplot`, `scatterplot` and so on.
- Scatter plot for Bland-Altman
- Box plot for data distribution
- Scatter Plot for Regression
- ROC plot
- Line plot for by visit

Adds functions to calculate the sample size for quantitative and qualitative trials, following the NMPA guidance's recommendation.
- Test for One Proportion
- Confidence Interval for One Proportion
- Pearson's Correlation Tests
- Confidence Interval for Pearson's Correlation

## Question

For the choice of class, either S3 nor S4. 
Considering the `mcr` package uses the S4 class `MCResult`, we'd better use the same structure as it. But the IVD's trial data is very simple, as shown below.

```{r}
head(mcr::creatinine)
```

Maybe some additional clinical data are required, such as age, sex, race, diagnostic and sample type. But the measurements are more critical, and they are the main data should be used in each model and method.

So we don't need to create a specific class for store trial data, just use `data.frame` or `tibble()` is fine for input data.

It appears that we have some outputs from various methodologies, like regression, Bland-Altman, and so on. I can not think of a general class that would handle these methods. So I'm not sure that maybe S3 is a simplified way to implement?


### Validation

If we choose S3 not S4, it appears that we don't need the validation process.

